Viscous Erosion in Porous Media

### 2018-01-11 docs/methods/erosion_methods.tex
started to rearrange the paper

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Layer Potentials for Quantities of Interest}
\label{sec:qois}
There are several quantities besides the velocity that need to be
computed.  First, the governing equations~\eqref{eqn:erosionModel}
require the shear stress along $\gamma_\ell$.  In addition, to study
characteristics of the flow, the vorticity, pressure, and drag, and
resistivity will also be computed.  Once
equation~\eqref{eqn:completed_DLP} is solved for the density function,
Stokeslets, and rotlets, all of these quantities can be computed by
evaluating layer potentials that we now derive.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Shear Stress}
The shear stress is
\begin{align*}
  \tau = -\left(\nabla \uu + \nabla \uu^T \right)\nn \cdot \ss,
\end{align*}
where $\ss$ and $\nn$ are the unit tangent and normal vectors,
respectively (Figure~\ref{fig:schematic}).  The velocity field of the
completed double-layer potential~\eqref{eqn:completed_DLP} involves
three terms: the double-layer potential $\DD$, the Stokeslets $S$, and
the rotlets $R$.  We first compute the deformation tensor $\ssigma =
\frac{1}{2}(\nabla\uu + \nabla\uu^T)$ for each of these terms
individually.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Vorticity}
For the Stokes equations, on boundaries that have a no-slip boundary
condition, the shear stress is equivalent to the vorticity $w = v_x -
u_y$.  Therefore, to compute $w(\xx)$ for $\xx \in \overline{\Omega}$,
we only need to compute $w(\xx)$ for $\xx \in \Omega$.  As we did for
the shear stress, we compute the vorticity due to the double-layer
potential, Stokeslets, and rotlets individually.  The vorticity of the
double-layer potential~\eqref{eqn:velocityDLP} at $\xx \in \Omega$ is
\begin{align}
  w^{\DD}(\xx) = -\frac{1}{\pi}\int_{\bd\Omega} 
    \frac{(\rr \cdot \nn^\perp) + (\rr \cdot \nn)}{\rho^4}
    (\rr \cdot \eeta) ds_{\yy}.
  \label{eqn:vorticityDLP}
\end{align}
The vorticity due to the Stokeslet is
\begin{align*}
  w^S(\xx) = -\frac{1}{\pi} \sum_{\ell=1}^{M} 
    \frac{\rr \cdot \llambda_\ell^\perp}{\rho^2},
\end{align*}
and the vorticity of a rotlet is zero.  Therefore, the vorticity at $\xx
\in \Omega$ is
\begin{align*}
  w(\xx) = w^\DD(\xx) + w^S(\xx).
\end{align*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Pressure}
To compute the pressure at $\xx \in \overline{\Omega}$, we follow the
same procedure used for the shear stress.  We start by computing the
pressure of the double-layer potential for $\xx \in \Omega$ and add the
appropriate jump if $\xx \in \Gamma$. For $\xx \in \Omega$, the pressure
is
\begin{align}
  p^{D}(\xx) = -\frac{1}{\pi}\int_{\bd\Omega} \frac{1}{\rho^2}
    \left(I - 2 \frac{\rr \otimes \rr}{\rho^2}\right) 
    \nn \cdot \eeta(\yy) \,ds_\yy.
    \label{eqn:pressureDLP}
\end{align}
To compute the pressure at $\xx_0 \in \bd\Omega$, we consider the limit
as $\xx$ tends to $\xx_0$.  The limiting value is given
by~\cite{qua-bir2014a}
\begin{align}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}p^\DD(\xx) =  
    \pderiv{\eeta}{\ss} \cdot \ss + p^{\DD}(\xx_0).
  \label{eqn:pressureJump}
\end{align}
The pressure due to the Stokeslets is
\begin{align}
  p^S(\xx_0) = \sum_{\ell=1}^{M}\frac{\rr \cdot \llambda_\ell}{2\pi\rho^2},
  \label{eqn:shearPressureSR}
\end{align}
and the pressure due to the rotlets is zero.  Therefore, the pressure is
\begin{align*}
  p(\xx) &= p^{\DD}(\xx) + p^{S}(\xx), \quad &&\xx \in \Omega,\\
  p(\xx) &= \pderiv{\eeta}{\ss} \cdot \ss + p^{\DD}(\xx) + 
              p^{S}(\xx), && \xx \in \bd\Omega.
\end{align*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Drag}
\todo[inline]{check all signs in this section.  Fortran code assumes a
normal pointing into the eroding body and a tangent vector that
corresponds to a couterclockwise parameterization}
With the pressure and shear stress, the drag on a single body with
boundary $\gamma_\ell$ is
\begin{align}
  F_D = \int_{\gamma_\ell} (-p \nn + \tau \ss)\,ds.
  \label{eqn:drag}
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Resistivity}
\begin{align}
  some equation
  \label{eqn:resistivity}
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Quadrature: On Boundary Calculations} 
\label{sec:on_boundary}
We start by discussing the quadrature methods that are used to compute
the velocity~\eqref{eqn:DLP}, shear stress~\eqref{eqn:shearStressDLP},
pressure~\eqref{eqn:pressureDLP}, and drag~\eqref{eqn:drag} at target
points on the boundary of the domain.  The velocity on $\bd\Omega$ is
required to solve the integral equation~\eqref{eqn:completed_BIE} for
the density function, Stokeslets and rotlets.  Once these are computed,
computing the shear stress, pressure, and drag is a post-processing
step.  All these integrals involve integrands that are periodic.  For
smooth integrands, the trapezoid rule guarantees spectral
accuracy~\cite{tre-wei2014}.  However, some of the integrands contain a
singularity whose strength will be reduced with singularity subtraction,
and then spectral accuracy is achieved with odd-even
integration~\cite{sid-isr1988} must be used.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Velocity}
To numerically evaluate the completed double-layer
potential~\eqref{eqn:completed_BIE}, we le $N = MN_\iin + N_\out$ be the
total number of discretization points of $\bd\Omega$.  The double-layer
potential in~\eqref{eqn:DLP} is discretized with the trapezoid rule as
\begin{align}
  \DD[\eeta](\xx_i) = \sum_{j=1}^{N} K(\xx_i,\xx_j) \eeta(\xx_j) 
      \Delta s_{j}, \quad i=1,\ldots,N,
  \label{eqn:trapDLP}
\end{align}
where $\Delta s_j$ is the arclength term of $\bd\Omega$ at
$\xx_j$, and
\begin{align*}
  K(\xx,\yy) = \frac{1}{\pi} \frac{\rr \cdot \nn}{\rho^2} 
      \frac{\rr \otimes \rr}{\rho^2},
\end{align*}
where $\rr = \xx - \yy$ and $\rho = \|\rr\|$, is the kernel of the
double-layer potential.  The diagonal term of~\eqref{eqn:trapDLP} is
replaced with the limiting value of the double-layer potential
\begin{align*}
  D(\xx,\xx) = -\frac{\kappa(\xx)}{2\pi}(\ss(\xx) \otimes \ss(\xx)),
\end{align*}
where $\kappa(\xx)$ is the curvature and $\ss(\xx)$ is the unit tangent
vector at $\xx \in \bd\Omega$.  The strengths of the
Stokeslets~\eqref{eqn:stokeslet} and the rotlets~\eqref{eqn:rotlet} are
also discretized with the trapezoid rule.  Applying this discretization
strategy to~\eqref{eqn:completed_BIE}, the dense linear system that
needs to be solved is
\begin{equation}
  \label{eqn:completed_BIE_discretized}
  \begin{aligned}
    \ff(\xx_i) &= -\frac{1}{2}\ssigma(\xx_i) + \sum_{j=1}^{N} 
      K(\xx_i,\xx_j) \eeta(\xx_j) \Delta s_{j} + 
      \sum_{\ell=1}^M S[\llambda_\ell,\cc_\ell](\xx_i) \\
      &\hspace{40pt}
      +\sum_{\ell=1}^M R[\xi_\ell,\cc_\ell](\xx_i) 
      +\sum_{\xx_j \in \Gamma} (\nn(\xx_i) \otimes
      \nn(\xx_j))\eeta(\xx_j), \\
    \llambda_\ell &= \sum_{\xx_i \in \gamma_\ell} \eeta(\xx_i) 
      \Delta s_i, \\ 
    \mu_\ell &= \sum_{\xx_i \in \gamma_\ell}
      {(\xx_i - \cc_\ell)}^\perp \cdot \eeta(\xx_i) \Delta s_i,
  \end{aligned}
\end{equation}
for $i=1,\ldots,N$, $\ell=1,\ldots,M$.  Since all the integrands are
smooth and periodic, by using the trapezoid rule, we are guaranteed that
the numerical solutions converges to the true solution with spectral
accuracy.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Shear Stress}
We have decomposed the deformation tensor into the contributions due to
three different terms.  The contributions from the Stokeslets and the
rotlets can be directly evaluated since $\rr \neq 0$ for all $\xx \in
\gamma$.  However, the kernel of the deformation tensor due to the
double-layer potential~\eqref{eqn:shearStressDLP} 
\begin{align*}
  K_\ssigma(\xx,\yy)\eeta(\yy) &= \frac{1}{2\pi} \left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn)
    \right. \\ & \hspace{25pt}+ \left.
    \frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
  \right),
\end{align*}
has a singularity that is asymptotic to $\bigO(\rho^{-2})$ when $\yy
\rightarrow \xx$.  The singularity strength can be reduced since the
velocity corresponding to a constant density function is
constant~\cite{poz1992}.  Therefore, the deformation tensor of the
double-layer potential is equivalent to
\begin{align*}
  \ssigma^{\DD}(\xx) = \int_{\bd\Omega}K_\ssigma(\xx,\yy)
      (\eeta(\yy) - \eeta(\xx)) \, ds_{\yy}, \quad \xx \in \bd\Omega.
\end{align*}
This expression for the deformation tensor has a singularity that is
asymptotic to $\bigO(\rho^{-1})$ when $\yy \rightarrow \xx$.  Finally,
to achieve spectral accuracy, we apply odd-even
integration~\cite{sid-isr1988}.  That is, the deformation tensor due to
the double-layer potential at a discretization point $\xx_i$ on body
$\gamma_\ell$, with $i$ being odd, is approximated as
\begin{align}
  \ssigma^{\DD}(\xx_i) = \sum_{\xx_j \notin \gamma_\ell}
    K_\ssigma(\xx_i,\xx_j) \eeta(\xx_j) \Delta s_j + 
  \sum_{\substack{\xx_j \in \gamma_\ell \\ j \: \mathtt{even}}}
    K_\ssigma(\xx_i,\xx_j) (\eeta(\xx_j) - \eeta(\xx_i)) \Delta s_j,
  \label{eqn:stressOddEven}
\end{align}
and a similar approximation is used when $i$ is even.  The deformation
tensor due to the Stokeslets and rotlets is given in
equation~\eqref{eqn:shearStressSR}.  Finally, the jump in the
deformation tensor~\eqref{eqn:shearStressJump} is added, and the
required arclength derivative is computed in Fourier space.  Then, the
shear stress can be computed by computing the deformation tensor in the
normal vector, and taking the inner product with the tangent vector.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Pressure}
Similar to the shear stress, the singularity of the pressure
corresponding to the double-layer potential~\eqref{eqn:pressureDLP} is
too strong for the trapezoid rule to be directly applied.  Therefore, we
again subtract a constant density function and write the pressure as
\begin{align*}
  p^\DD(\xx) = -\frac{1}{\pi}\int_{\bd\Omega} \frac{1}{\rho^2}
    \left(I - 2 \frac{\rr \otimes \rr}{\rho^2}\right) 
    \nn \cdot (\eeta(\yy) - \eeta(\xx)) \,ds_\yy, 
    \quad \xx \in \bd\Omega
\end{align*}
Then, identical to how the shear stress is
computed~\eqref{eqn:stressOddEven}, odd-even integration is used to
evaluate this layer potential for pressure.  The pressure due to the
Stokeslets are given in equation~\eqref{eqn:shearPressureSR}.  Finally,
to add the jump condition~\eqref{eqn:pressureJump}, Fourier
differentiation is used to compute the arclength derivative.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Drag}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Near-singular integration}
To accurately compute quantities such as the velocity at points $\xx \in
\Omega$, we adopt a near-singular integration scheme entirely based on
upsampling.  Since we are only interested in computing the velocity and
vorticity for visualizing the flow, and the pressure is only required
far away from any boundary, only a finite amount of upsampling is
required to achieve an acceptable error.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Resistivity}

@@

%% TIME-STEPPING %%
\subsection{Time-stepping with the {\thL} method}

With the shear stress computed, we are now ready to evolve the interface of each body. As described in Section \ref{sec:thetaL}, we use the {\thL} framework, with $\theta$ being the tangent angle and $L$ the total arc length of a given boundary, and with the boundaries parameterized by normalized arc length $\alpha = 2 \pi s / L$.

In addition to the curvature-dependent term in \eqref{eqn:Vn}, we regularize boundary evolution by applying a narrow Gaussian filter to absolute shear stress
\begin{equation}
\atausig = \int_{\gamma} \frac{1}{\sqrt{2\pi} \sigma} \exp \left(- \frac{\alpha'^2}{2 \sigma^2}\right) \abs{ \tau(\alpha - \alpha') } \, d\alpha'
\end{equation}
Here $\sigma$ is the small parameter that determines the width of the filter. The corresponding normal and tangential interface velocities are the given by
\begin{align}
& \Vnsig = \atausig +  \eps \mean{\atausig}   \left(\pderiv{\theta}{\alpha} - 1 \right) \\
& \pderiv{\Vssig}{\alpha} = \thalpha \Vnsig - \mean{\thalpha \Vnsig}
\end{align}

%The interface evolution is then given by
%\begin{equation}
%\dot{\xx}(t) = \Vn \nn + \Vs \ss.
%\end{equation}

For simplicity, we introduce the variables
\begin{align}
\label{eq:M}
& \Mterm = - 2 \pi \mean{\thalpha \Vnsig} \\
\label{eq:N}
& \NLterm = \frac{2 \pi}{L} \left( \pderiv{\atausig}{\alpha} + \thalpha \Vssig \right) \\
\label{eq:elfun}
& \elfun = \frac{2 \pi  \mean{\atausig}}{L}
\end{align}
So that evolution equations \eqref{eqn:Levo1} and \eqref{eqn:thetaevo1} can be compactly expressed as
\begin{align}
\label{eqn:Levo2}
& \tderiv{L}{t} = \Mterm \\
\label{eqn:thetaevo2}
& \pderiv{\theta}{t} = \eps \elfun \ppd{\theta}{\alpha} + \NLterm
\end{align}

Here $\NLterm$ represents the nonlinear terms (and nonlocal etc) in the PDE for $\theta$.
We will evolve $\theta$ in spectral space, and so introduce the Fourier series
\begin{align}
& \theta(\alpha,t) = \alpha + \FourierSum \thhat_k e^{ i k \alpha} \\
& \NLterm(\alpha,t)  = \FourierSum \hat{\NLterm}_k e^{ i k \alpha}
\end{align}
The careful reader will notice that we have separated out the linear part of $\theta$, so that the remaining function is smooth and has a rapidly decaying Fourier series.
In spectral space, \eqref{eqn:thetaevo2} produces the system of ODEs
\begin{equation}
\tderiv{\thhat_k}{t} +  \eps k^2  \elfun \thhat_k = \hat{\NLterm}_k
\end{equation}
Each ODE is first-order and linear (though variable-coefficient due to $\elfun$) and thus can be solved by using the integrating factor
\begin{equation}
\label{eq:mu}
\mu_k = \exp \left( \eps k^2 \int \elfun \, dt \right)
\end{equation}
We are therefore led to the coupled system of ODEs
\begin{align}
\label{eqn:Levo2}
& \tderiv{L}{t} = \Mterm \\
\label{eqn:thetaevo2}
& \tderiv{}{t}\left( \mu_k \thhat_k \right) = \mu_k \hat{\NLterm}_k
\end{align}
We solve system \eqref{eqn:Levo2}--\eqref{eqn:thetaevo2} with an explicit, second-order Runge-Kutta method (RK2). In particular, we choose the {\em midpoint} RK2 method so as to maximize the smoothing effect of the curvature term.

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Computing the Shear Stress}
The shear stress is
\begin{align*}
  \tau = -\left(\nabla \uu + \nabla \uu^T \right)\nn \cdot \ss,
\end{align*}
where $\ss$ and $\nn$ are the unit tangent and normal vectors,
respectively (Figure~\ref{fig:schematic}).  The velocity field of the
completed double-layer potential~\eqref{eqn:completed_DLP} involves
three terms: the double-layer potential $\DD$, the Stokeslets $S$, and
the rotlets $R$.  We first compute the deformation tensor $\ssigma =
\frac{1}{2}(\nabla\uu + \nabla\uu^T)$ for each of these terms
individually.

We have decomposed the deformation tensor into the contributions due to
three different terms.  The contributions from the Stokeslets and the
rotlets can be directly evaluated since $\rr \neq 0$ for all $\xx \in
\gamma$.  However, the kernel of the deformation tensor due to the
double-layer potential~\eqref{eqn:shearStressDLP} 
\begin{align*}
  K_\ssigma(\xx,\yy)\eeta(\yy) &= \frac{1}{2\pi} \left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn)
    \right. \\ & \hspace{25pt}+ \left.
    \frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
  \right),
\end{align*}
has a singularity that is asymptotic to $\bigO(\rho^{-2})$ when $\yy
\rightarrow \xx$.  The singularity strength can be reduced since the
velocity corresponding to a constant density function is
constant~\cite{poz1992}.  Therefore, the deformation tensor of the
double-layer potential is equivalent to
\begin{align*}
  \ssigma^{\DD}(\xx) = \int_{\bd\Omega}K_\ssigma(\xx,\yy)
      (\eeta(\yy) - \eeta(\xx)) \, ds_{\yy}, \quad \xx \in \bd\Omega.
\end{align*}
This expression for the deformation tensor has a singularity that is
asymptotic to $\bigO(\rho^{-1})$ when $\yy \rightarrow \xx$.  Finally,
to achieve spectral accuracy, we apply odd-even
integration~\cite{sid-isr1988}.  That is, the deformation tensor due to
the double-layer potential at a discretization point $\xx_i$ on body
$\gamma_\ell$, with $i$ being odd, is approximated as
\todo[inline]{fix up this section and move numerics to the next section}
\begin{align}
  \ssigma^{\DD}(\xx_i) = \sum_{\xx_j \notin \gamma_\ell}
    K_\ssigma(\xx_i,\xx_j) \eeta(\xx_j) \Delta s_j + 
  \sum_{\substack{\xx_j \in \gamma_\ell \\ j \: \mathtt{even}}}
    K_\ssigma(\xx_i,\xx_j) (\eeta(\xx_j) - \eeta(\xx_i)) \Delta s_j,
  \label{eqn:stressOddEven}
\end{align}
and a similar approximation is used when $i$ is even.  The deformation
tensor due to the Stokeslets and rotlets is given in
equation~\eqref{eqn:shearStressSR}.  Finally, the jump in the
deformation tensor~\eqref{eqn:shearStressJump} is added, and the
required arclength derivative is computed in Fourier space.  Then, the
shear stress can be computed by computing the deformation tensor in the
normal vector, and taking the inner product with the tangent vector.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We start by discussing the quadrature methods that are used to compute
the velocity~\eqref{eqn:DLP} and shear stress~\eqref{eqn:shearStressDLP}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Solving the BIE}
To numerically evaluate the completed double-layer
potential~\eqref{eqn:completed_BIE}, we le $N = MN_\iin + N_\out$ be the
total number of discretization points of $\bd\Omega$.  The double-layer
potential in~\eqref{eqn:DLP} is discretized with the trapezoid rule as
\begin{align}
  \DD[\eeta](\xx_i) = \sum_{j=1}^{N} K(\xx_i,\xx_j) \eeta(\xx_j) 
      \Delta s_{j}, \quad i=1,\ldots,N,
  \label{eqn:trapDLP}
\end{align}
where $\Delta s_j$ is the arclength term of $\bd\Omega$ at
$\xx_j$, and
\begin{align*}
  K(\xx,\yy) = \frac{1}{\pi} \frac{\rr \cdot \nn}{\rho^2} 
      \frac{\rr \otimes \rr}{\rho^2},
\end{align*}
where $\rr = \xx - \yy$ and $\rho = \|\rr\|$, is the kernel of the
double-layer potential.  The diagonal term of~\eqref{eqn:trapDLP} is
replaced with the limiting value of the double-layer potential
\begin{align*}
  D(\xx,\xx) = -\frac{\kappa(\xx)}{2\pi}(\ss(\xx) \otimes \ss(\xx)),
\end{align*}
where $\kappa(\xx)$ is the curvature and $\ss(\xx)$ is the unit tangent
vector at $\xx \in \bd\Omega$.  The strengths of the
Stokeslets~\eqref{eqn:stokeslet} and the rotlets~\eqref{eqn:rotlet} are
also discretized with the trapezoid rule.  Applying this discretization
strategy to~\eqref{eqn:completed_BIE}, the dense linear system that
needs to be solved is
\begin{equation}
  \label{eqn:completed_BIE_discretized}
  \begin{aligned}
    \ff(\xx_i) &= -\frac{1}{2}\ssigma(\xx_i) + \sum_{j=1}^{N} 
      K(\xx_i,\xx_j) \eeta(\xx_j) \Delta s_{j} + 
      \sum_{\ell=1}^M S[\llambda_\ell,\cc_\ell](\xx_i) \\
      &\hspace{40pt}
      +\sum_{\ell=1}^M R[\xi_\ell,\cc_\ell](\xx_i) 
      +\sum_{\xx_j \in \Gamma} (\nn(\xx_i) \otimes
      \nn(\xx_j))\eeta(\xx_j), \\
    \llambda_\ell &= \sum_{\xx_i \in \gamma_\ell} \eeta(\xx_i) 
      \Delta s_i, \\ 
    \mu_\ell &= \sum_{\xx_i \in \gamma_\ell}
      {(\xx_i - \cc_\ell)}^\perp \cdot \eeta(\xx_i) \Delta s_i,
  \end{aligned}
\end{equation}
for $i=1,\ldots,N$, $\ell=1,\ldots,M$.  Since all the integrands are
smooth and periodic, by using the trapezoid rule, we are guaranteed that
the numerical solutions converges to the true solution with spectral
accuracy.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Shear Stress}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Near-singular integration}
To accurately compute quantities such as the velocity at points $\xx \in
\Omega$, we adopt a near-singular integration scheme entirely based on
upsampling.  Since we are only interested in computing the velocity and
vorticity for visualizing the flow, and the pressure is only required
far away from any boundary, only a finite amount of upsampling is
required to achieve an acceptable error.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% TIME-STEPPING %%
\subsection{Time-stepping with the {\thL} method}

With the shear stress computed, we are now ready to evolve the interface
of each body. As described in Section~\ref{sec:thetaL}, we use the
{\thL} framework, with $\theta$ being the tangent angle and $L$ the
total arc length of a given boundary, and with the boundaries
parameterized by normalized arc length $\alpha = 2 \pi s / L$.

In addition to the curvature-dependent term in~\eqref{eqn:Vn}, we
regularize boundary evolution by applying a narrow Gaussian filter to
absolute shear stress
\begin{equation}
\atausig = \int_{\gamma} \frac{1}{\sqrt{2\pi} \sigma} \exp \left(- \frac{\alpha'^2}{2 \sigma^2}\right) \abs{ \tau(\alpha - \alpha') } \, d\alpha'
\end{equation}
Here $\sigma$ is the small parameter that determines the width of the
filter. The corresponding normal and tangential interface velocities are
the given by
\begin{align}
& \Vnsig = \atausig +  \eps \mean{\atausig}   \left(\pderiv{\theta}{\alpha} - 1 \right) \\
& \pderiv{\Vssig}{\alpha} = \thalpha \Vnsig - \mean{\thalpha \Vnsig}
\end{align}

For simplicity, we introduce the variables
\begin{align}
\label{eq:M}
& \Mterm = - 2 \pi \mean{\thalpha \Vnsig} \\
\label{eq:N}
& \NLterm = \frac{2 \pi}{L} \left( \pderiv{\atausig}{\alpha} + \thalpha \Vssig \right) \\
\label{eq:elfun}
& \elfun = \frac{2 \pi  \mean{\atausig}}{L}
\end{align}
So that evolution equations`\eqref{eqn:Levo1} and~\eqref{eqn:thetaevo1}
can be compactly expressed as
\begin{align}
\label{eqn:Levo2}
& \tderiv{L}{t} = \Mterm \\
\label{eqn:thetaevo2}
& \pderiv{\theta}{t} = \eps \elfun \ppd{\theta}{\alpha} + \NLterm
\end{align}

Here $\NLterm$ represents the nonlinear terms (and nonlocal etc) in the
PDE for $\theta$.  We will evolve $\theta$ in spectral space, and so
introduce the Fourier series
\begin{align}
& \theta(\alpha,t) = \alpha + \FourierSum \thhat_k e^{ i k \alpha} \\
& \NLterm(\alpha,t)  = \FourierSum \hat{\NLterm}_k e^{ i k \alpha}
\end{align}
The careful reader will notice that we have separated out the linear
part of $\theta$, so that the remaining function is smooth and has a
rapidly decaying Fourier series.  In spectral
space,~\eqref{eqn:thetaevo2} produces the system of ODEs
\begin{equation}
\tderiv{\thhat_k}{t} +  \eps k^2  \elfun \thhat_k = \hat{\NLterm}_k
\end{equation}
Each ODE is first-order and linear (though variable-coefficient due to
$\elfun$) and thus can be solved by using the integrating factor
\begin{equation}
\label{eq:mu}
\mu_k = \exp \left( \eps k^2 \int \elfun \, dt \right)
\end{equation}
We are therefore led to the coupled system of ODEs
\begin{align}
\label{eqn:Levo2}
& \tderiv{L}{t} = \Mterm \\
\label{eqn:thetaevo2}
& \tderiv{}{t}\left( \mu_k \thhat_k \right) = \mu_k \hat{\NLterm}_k
\end{align}
We solve system~\eqref{eqn:Levo2}--\eqref{eqn:thetaevo2} with an
explicit, second-order Runge-Kutta method (RK2). In particular, we
choose the {\em midpoint} RK2 method so as to maximize the smoothing
effect of the curvature term.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Post Processing Quantities of Interest\label{s:qoi}}

There are several quantities besides the velocity that need to be
computed.  First, the governing equations~\eqref{eqn:erosionModel}
require the shear stress along $\gamma_\ell$.  In addition, to study
characteristics of the flow, the vorticity, pressure, and drag, and
resistivity will also be computed.  Once
equation~\eqref{eqn:completed_DLP} is solved for the density function,
Stokeslets, and rotlets, all of these quantities can be computed by
evaluating layer potentials that we now derive.

pressure~\eqref{eqn:pressureDLP}, and drag~\eqref{eqn:drag} at target
points on the boundary of the domain.  The velocity on $\bd\Omega$ is
required to solve the integral equation~\eqref{eqn:completed_BIE} for
the density function, Stokeslets and rotlets.  Once these are computed,
computing the shear stress, pressure, and drag is a post-processing
step.  All these integrals involve integrands that are periodic.  For
smooth integrands, the trapezoid rule guarantees spectral
accuracy~\cite{tre-wei2014}.  However, some of the integrands contain a
singularity whose strength will be reduced with singularity subtraction,
and then spectral accuracy is achieved with odd-even
integration~\cite{sid-isr1988} must be used.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Vorticity}
For the Stokes equations, on boundaries that have a no-slip boundary
condition, the shear stress is equivalent to the vorticity $w = v_x -
u_y$.  Therefore, to compute $w(\xx)$ for $\xx \in \overline{\Omega}$,
we only need to compute $w(\xx)$ for $\xx \in \Omega$.  As we did for
the shear stress, we compute the vorticity due to the double-layer
potential, Stokeslets, and rotlets individually.  The vorticity of the
double-layer potential~\eqref{eqn:velocityDLP} at $\xx \in \Omega$ is
\begin{align}
  w^{\DD}(\xx) = -\frac{1}{\pi}\int_{\bd\Omega} 
    \frac{(\rr \cdot \nn^\perp) + (\rr \cdot \nn)}{\rho^4}
    (\rr \cdot \eeta) ds_{\yy}.
  \label{eqn:vorticityDLP}
\end{align}
The vorticity due to the Stokeslet is
\begin{align*}
  w^S(\xx) = -\frac{1}{\pi} \sum_{\ell=1}^{M} 
    \frac{\rr \cdot \llambda_\ell^\perp}{\rho^2},
\end{align*}
and the vorticity of a rotlet is zero.  Therefore, the vorticity at $\xx
\in \Omega$ is
\begin{align*}
  w(\xx) = w^\DD(\xx) + w^S(\xx).
\end{align*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Pressure}
To compute the pressure at $\xx \in \overline{\Omega}$, we follow the
same procedure used for the shear stress.  We start by computing the
pressure of the double-layer potential for $\xx \in \Omega$ and add the
appropriate jump if $\xx \in \Gamma$. For $\xx \in \Omega$, the pressure
is
\begin{align}
  p^{D}(\xx) = -\frac{1}{\pi}\int_{\bd\Omega} \frac{1}{\rho^2}
    \left(I - 2 \frac{\rr \otimes \rr}{\rho^2}\right) 
    \nn \cdot \eeta(\yy) \,ds_\yy.
    \label{eqn:pressureDLP}
\end{align}
To compute the pressure at $\xx_0 \in \bd\Omega$, we consider the limit
as $\xx$ tends to $\xx_0$.  The limiting value is given
by~\cite{qua-bir2014a}
\begin{align}
  \lim_{\substack{\xx \rightarrow \xx_0 \\ \xx \in \Omega}}p^\DD(\xx) =  
    \pderiv{\eeta}{\ss} \cdot \ss + p^{\DD}(\xx_0).
  \label{eqn:pressureJump}
\end{align}
The pressure due to the Stokeslets is
\begin{align}
  p^S(\xx_0) = \sum_{\ell=1}^{M}\frac{\rr \cdot \llambda_\ell}{2\pi\rho^2},
  \label{eqn:shearPressureSR}
\end{align}
and the pressure due to the rotlets is zero.  Therefore, the pressure is
\begin{align*}
  p(\xx) &= p^{\DD}(\xx) + p^{S}(\xx), \quad &&\xx \in \Omega,\\
  p(\xx) &= \pderiv{\eeta}{\ss} \cdot \ss + p^{\DD}(\xx) + 
              p^{S}(\xx), && \xx \in \bd\Omega.
\end{align*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Drag}
\todo[inline]{check all signs in this section.  Fortran code assumes a
normal pointing into the eroding body and a tangent vector that
corresponds to a couterclockwise parameterization}
With the pressure and shear stress, the drag on a single body with
boundary $\gamma_\ell$ is
\begin{align}
  F_D = \int_{\gamma_\ell} (-p \nn + \tau \ss)\,ds.
  \label{eqn:drag}
\end{align}
